Profiles of Wind Veer and Wind Speed in Turbulent Ekman Flow

### 2024-06-27 write1/manuscript.tex
update on part 2

\documentclass[a4paper,11pt]{amsart}

\usepackage[margin=3.5cm]{geometry}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[footnotesize]{caption}
\usepackage{xcolor}
\usepackage{dirtytalk}
\usepackage{colortbl}
\usepackage{booktabs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BIBLIOGRAPHY DETAILS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[maxcitenames=2,style=authoryear-comp,firstinits=true,citestyle=authoryear,natbib=true,backend=bibtex]{biblatex}
\DeclareNameAlias{sortname}{family-given}
\renewcommand*{\multinamedelim}{\addcomma\space}
\renewcommand*{\finalnamedelim}{~and\space}
\setlength{\bibitemsep}{\baselineskip}
\setlength\parindent{3em}
\setlength\parskip{0.5em}
\renewbibmacro{in:}{}

\AtEveryBibitem{%
  \clearfield{url}
  \clearfield{urlday}
  \clearfield{urlmonth}
  \clearfield{urlyear}
  \clearfield{issn}
  \clearfield{month}
%%%  \clearfield{doi}%
  \clearfield{keywords}
  \clearfield{note}
}

\bibliography{ansorge}

@@

\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
\newcommand{\quotation}[1]{"#1"}

% TO BE USED IN MATH-ENVIRONMENT 
\newcommand{\SFC}{\mathrm{sfc}}
\newcommand{\p}{\partial}
\newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}
\newcommand{\RO}{\mathrm{Ro}}  
\newcommand{\RE}{\mathrm{Re}}
\newcommand{\LR}{\Lambda_\RO}

@@

Wind veer away from its geostrophic direction is
(i) due to direct frictional effects in the very vicinity of the surface and
(ii) due to turbulence which exerts indirect frictional effects; these effects
cause a slow-down of the mean wind reducing the Coriolis force thus turning the wind
in favor of the pressure gradient force. 
%%
Not only does the veering set the frame of reference for surface layer theory, it also has
effects at small and large scales from large-scale dispersion via plume spreading to
cyclone spin-down \citep{svensson:BM2009}
and on the capabilities of data assimilation and accuracy of surface flux estimates \citep{brown:QJR2005}. 
%
From a large-scale perspective, the veering of wind across the planetary boundary layer determines
the amount of cross-isobaric mass-flux, commonly referred to as 'Ekman pumping' \citep{ekman:AMA1905},
and it is thus a key factor in the life-cycle of large-scale synoptic systems.
%
% %% This shear is probably not related to Ekman-turning 
% On the mesoscale, shear in the surface layer a crucial factor for the initiation and sustainment of strong convective events (REFERENCE?).
% 
Within the atmospheric boundary layer (ABL), directional shear of the wind in
the upper part of the surface layer may cause a systmatic yaw for tall wind power generation devices
where blades reach into the Ekman layer, i.e. that part of the boundary layer where the wind starts to turn;
an exact estimate of such effects is critical in the site assessments for wind farms
\citep{calaf:PF2010, mirocha:WES2018}. 
%
\par
%
In the planetary boundary layer, wind veer is characterized by the surface veering angle $\alpha$
defined as the angle between the negative surface shear stress $\tau_\SFC$ and the geostrophic wind.
%
Surface veering $\alpha$ and geostrophic drag $Z\equiv u_\star/G$, where $u_\star\equiv\sqrt{\nu|\tau_\SFC|}$,
uniquely determine the surface drag $\tau_\SFC$ in a turbulent Ekman flow. 
%
In any quantitative description of the surface layer, the friction velocity $u_\star$ is the dynamic scale
and $\alpha$ defines the alignment of the frame of reference.
%
Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer, 
and \citet{rossby:PPO1935} constrained the two parameters based on integral relations in the ABL.
%
Asymptotic similarity theory was later used by \citet{tennekes:JAS1973, blackadar:JAS1968},
and--based on his seminal direct numerical simulations (DNS) of Ekman flow--,~\cite{spalart:JFM1989} suggested
a modification to take into account effects of low to intermediate Reynolds numbers.
Later on, constants were re-evaluated with a focus on the atmospheric boundary layer based on
observations \citep{hogstrom:BM1988,hogstrom:BM1996}
and numerical modelling \citep{spalart:PF2008, spalart:PF2009, ansorge:BM2014, ansorge:BM2019}.
\todo{Refer to Pena et al, 2010} 
%
\par
%
Attempts were also undertaken to obtain profiles of the wind speed: 
One approach is to match the inner and outer layer at a reference height;
\citet{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height to match
the wind speed profiles, which, however, requires to externally prescribe the veering $\alpha$ at that height. 
A one-dimensional profile with constant veering is given by \citet[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}.

\citet{gryning:BM2007} present an extension of the wind-speed profile beyond the surface layer
using a neutral referece profile and a stability correction;
\cite{kelly:BM2010}, based on a probabilistic representation of stratification,
develop a model for the long-term mean windspeed
in the atmospheric boundary layer and compare this with observation at different sites; 
\cite{kelly:WE2016} demonstrate the effect of such improved model for wind-energy applications.
In consideration of the large scale separation in geophysical flow, the rotation of the wind in the
surface layer is often assumed negligible, and above investigations merely focus on the wind speed; 
that means, the veering of the wind with height is not described and there is little knowledge on the
profile of the span-wise velocity component and the precise shape of the hodograph in the limit of a
truly neutral Ekman boundary layer.

@@

Ekman-layer models and two-layer models of the ABL take into account rotational effects at higher altitudes,
for instance when the wind speed needs to be evaluated at heights on the order of 100-200\,\mathrm{m},

@@

Still, in \citeyear{jiang:JAS2018}, \citeauthor{jiang:JAS2018}
recognized that the outer part of the Ekman boundary layer receives less attention in comparison
with the surface layer and commence in studying the neutral problem by Large-Eddy simulation (LES).
They focus on the wind speed and find an extended logarithmic layer when considering the wind speed instead of
the shear-aligned component, and they eventually demonstrate by means of an analytical model that this vertical
extension of the logarithmic layer may be explained by a transfer of stress to the spanwise velocity component where
it is assumed that the shear vector $\tau(z)$ and stress vectors $(\p_zU,\p_zV)$ are aligned. 
More recently, \citet{ghannam:QJR2021} treated the horizontally averaged momentum budget to show that departures from
shear-alignment in the vicinity of the surface result in an integral of the wind veer ($\alpha_M$ in their notation)
over the height to very high accuracy ($\int_{z_0}^{H} \sin\alpha_M$ in their notation; their Eq.~(16)).

@@

homogeneous, stationary atmospheric boundary layer over a flat surface under neutral stratification. 

@@

roughness or other physical complications encountered in the realgeophysical system.

@@

relies on sub-grid closures that often assume alignment of the stress with gradients -- a questionable
pre-requisite when the wind rotates with height.

@@

research foci of LES for the ABL over the past 50 years \citep{stoll:BM2020}. 

@@

of \say{\emph{fully developed turbulence}  believed to occur for a sufficiently large but finite Reynolds number}

@@

We consider here incompressible, turbulent Ekman flow, that is, the turbulent flow over a flat rotating plate, as
a physical model for the truly neutral atmospheric boundary layer (ABL).

@@

$\alpha(z)= \sphericalangle \left( \underline{G},\underline{U}(z)\right)$. 

@@

that is governed by the physical parameters
(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_2^2}$ (where $\mathbf{G}=(G_1,G_2,0)$ is the geostrophic wind vector),

@@

The scales used in defining $\RE_\Lambda$ loose their relevance in describing the turbulent state of the flow.
The friction Reynolds number

@@

$\RE_\tau$ is suitable to compare the turbulence scale separation also to other wall-bounded flow geometries
such as rotating or non-rotating channel, pipe, or Couette flow.

@@

where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, and $g_j=G_j/G$ are non-dimensionalized components
geostrophic wind (with $g_{1}^2+g_{2}^2=1$ by construction).
These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral directions $Oxy$, a no-slip--no-penetration boundary at $z=0$, and

@@

  \caption{Direct numerical simulation datasets used in this
    work \label{tab:sim-setup}}
  \begin{tabular}{c r r c c c c c}

@@

       $125\,000$ &    $500$& 1.08 & $2048\times2048\times192$&4.1& 4.1& 1.05\\

@@

       $281\,250$ & $   750$& 1.08 & $3072\times3072\times384$&5.6&11.2& 1.60\\

@@

    $1\,280\,000$ & $1\,600$& 0.54 & $3860\times7680\times960$&8.6& 4.3& 1.00\\ 

@@

    normalized boundary layer depth $\delta_{95}/\delta$

@@

  \begin{tabular}{r r r r r r r }

@@

Velocity profiles in inner and outer units (Fig.~\ref{fig:profiles}a,b) are in accordance
with previous work \citep{coleman:JFM1992,spalart:PF2008,spalart:PF2009,ansorge:BM2014,ansorge:BM2019}:

@@

geostrophic wind $G$) governs the inviscid, i.e. $\RE$-independent,
normalization.

@@

No collapse in inner units is found for the profiles of spanwise velocity $V^{\alpha_\star+}(z^+)$;

@@

    Circles in \textbf{(a)} and \textbf{(b)} denote the height
    $z^-=0.15$, diamonds in \textbf{(c)} and \textbf{(d)} are for $z^+=50$

@@

\end{figure} 
The organization of the flow is depicted in terms of the turbulence kinetic energy in Fig.~\ref{fig:slices}.
%
In vicinity of the wall, at $y^+\approx 10$, (Fig.~\ref{fig:slices}a), elongated streaks aligned with the
surface shear stress dominate.
%
Moving away from the wall, to $y^+\approx 150$ (well within the logarithmic region), the structures
are larger and more isotropic, but they are still largely aligned with the surface shear stress.
%
In the upper part of the outer layer, around $y^+\approx 1000$, no clear signature of the
surface veering direction is found, and intense TKE structures (bright yellow)
are organized on a large spatial scale with with weaker eddies (greenish structures) and quiescent regions
in between.

@@

This precludes a drag law wherewith we begin this section (\ref{sec:drag-law}).
Based on the scaling arguments put forward in Sec.~\ref{sec:scaling}, we then develop
a formulation of the wind vector in the Ekman layer (Sec~\ref{sec:ekman}) and separately
for the streamwise (Sec.~\ref{sec:streamwise}) and spanwise (Sec.~\ref{sec:spanwise}) components
in the overlap and inner regions. 

@@

@@

\cite{optis:BM2014}, for instance, define an ``effective geostrophic wind vector that has the same magnitude of the
observed surface geostrophic wind and is rotated by the angle $\alpha$ [their nomenclature]'' to overcome the

@@

The text-book solution for Ekman flow makes use of the physical boundary conditions (BCs) for the atmospheric
boundary layer (no-slip at the bottom and geostrophic wind in the free atmosphere) and a constant eddy

@@

        \end{array}\right)\right]\\
    z_\mathrm{ek} &=& \delta_\mathrm{ek} (y^- -s_\mathrm{ek}). 

@@

@@

the low-$\RE$ correection suggested by \cite{spalart:JFM1989}.

@@

@@

\subsection{} 

@@

\begin{figure}
  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
  \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
    Thick, solid lines show theory, dashed lines data from DNS. }
  \label{fig:hodograph} 
\end{figure}

\section{Comparison with other theories}

\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
%
\par%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
$\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
%
\begin{figure}
  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
  \caption{Total velocity and veering (in degrees) vs inner and outer height.
    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
\end{figure} 
%
\section{Conclusions}
%
Applications:
\begin{itemize}
\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
\item
\end{itemize}

>>>>

\RequirePackage{fix-cm} 
\documentclass[smallcondensed,final]{svjour3}
\smartqed
\usepackage{appendix}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{array}
\usepackage{longtable}
\usepackage{natbib}

% additional packages
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{booktabs} 
\usepackage{amssymb}
\usepackage{hyperref}
\setcitestyle{aysep={}}
\linenumbers

\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
\renewenvironment{#1}%
{\linenomath\csname old#1\endcsname}%
{\csname oldend#1\endcsname\endlinenomath}}% 
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
\patchAmsMathEnvironmentForLineno{#1}%
\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

@@

\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
\renewcommand{\quotation}[1]{"#1"}

% TO BE USED IN MATH-ENVIRONMENT 
\newcommand{\SFC}{\mathrm{sfc}}
\newcommand{\p}{\partial}
\newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}
\newcommand{\RO}{\mathrm{Ro}}  
\newcommand{\RE}{\mathrm{Re}}
\newcommand{\LR}{\Lambda_\RO} 
\newcommand{\DOI}[1]{\href{https://dx.doi.org/#1}{#1}}

\hypersetup{
  colorlinks=true,
  urlcolor=blue,
  citecolor=.,
  linkcolor=.
}

@@

The Coriolis force bends the path of motion on a rotating sphere and establishes geostrophic equilibrium when in balance with a pressure gradient force. Wind veer away from the wind direction in geostrophic equilibrium is
(i) due to direct frictional effects in the very vicinity of the surface and
(ii) due to turbulence which exerts indirect frictional effects; these effects
cause a slow-down of the mean wind reducing the Coriolis force thus turning the wind
in favor of the pressure gradient force. 
%%
Not only does the veering set the frame of reference for surface layer theory, it also has
effects at small and large scales from large-scale dispersion via plume spreading to
cyclone spin-down \citep{svensson:BM2009}
and on the capabilities of data assimilation and accuracy of surface flux estimates \citep{brown:QJR2005}. 
%
From a large-scale perspective, the veering of wind across the planetary boundary layer determines
the amount of cross-isobaric mass-flux, commonly referred to as 'Ekman pumping' \citep{ekman:AMA1905},
and it is thus a key factor in the life-cycle of large-scale synoptic systems.
%
% %% This shear is probably not related to Ekman-turning 
% On the meso-scale, shear in the surface layer a crucial factor for the initiation and sustained strong convective events (REFERENCE?).
% 
Within the atmospheric boundary layer (ABL), directional shear of the wind in
the upper part of the surface layer may cause a systematic yaw for tall wind power generation devices
where blades reach into the Ekman layer, i.e. that part of the boundary layer where the wind starts to turn;
an exact estimate of such effects is critical in the site assessments for wind farms
\citep{calaf:PF2010, mirocha:WES2018}. 
%
\par
%
In the planetary boundary layer, wind veer is characterized by the surface veering angle $\alpha$
defined as the angle between the negative surface shear stress $\tau_\SFC$ and the geostrophic wind.
%
Surface veering $\alpha$ and geostrophic drag $Z\equiv u_\star/G$, where the friction velocity $u_\star\equiv\sqrt{|\tau_\SFC|/\rho}$,
uniquely determine the surface drag $\tau_\SFC$ in a turbulent Ekman flow. 
%
In any quantitative description of the surface layer, the friction velocity $u_\star$ is the dynamic scale
and $\alpha$ defines the alignment of the frame of reference.
%
Knowledge about $u_\star$ and $\alpha$ is thus a prerequisite for any quantitative theory of the surface layer, 
and \citet{rossby:PPO1935} constrained the two parameters based on integral relations in the ABL.
%
Asymptotic similarity theory was later used by \cite{tennekes:JAS1973, blackadar:JAS1968},
and--based on his seminal direct numerical simulations (DNS) of Ekman flow--,~\cite{spalart:JFM1989} suggested
a modification to take into account effects of low to intermediate Reynolds numbers.
Later on, constants were re-evaluated with a focus on the ABL based on
observations \citep{hogstrom:BM1988,hogstrom:BM1996}
and numerical modelling \citep{spalart:PF2008, spalart:PF2009, ansorge:BM2014, ansorge:BM2019}.
%
\par
%
Attempts were also undertaken to obtain profiles of the wind speed: 
One approach is to match the inner and outer layer at a reference height;
\cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height $z_\text{Prandtl}$ to match
the wind speed profiles, which, however, requires external prescription of $\alpha(z_\text{prandtl})$, the veering at that height. 
A one-dimensional profile with constant veering is given by \citet[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}. 
%%%

\citet{gryning:BM2007} present an extension of the wind-speed profile beyond the surface layer
using a neutral reference profile and a stability correction;
\cite{kelly:BM2010}, based on a probabilistic representation of stratification,
develop a model for the long-term mean wind speed
in the ABL and compare this with observation at different sites; 
\cite{kelly:WE2016} demonstrate the effect of such improved model for wind-energy applications.
%
In consideration of the large scale separation in geophysical flow, the rotation of the wind in the
surface layer is often assumed negligible, and above investigations merely focus on the wind speed; 
that means, the veering of the wind with height is not described and there is little knowledge on the
profile of the span-wise velocity component and the precise shape of the hodograph in the limit of a
truly neutral Ekman boundary layer.
%
A climatology of wind turning in the ABL is given by \cite{lindvall:QJR2019} 
%
\cite{klein:PAM2021a} use a statistical turbulence modelling approach that yields a two-component velocity profile,
but they also find that the exact representation of turning is challenging.
%%%

@@

Ekman-layer models are roughly based on Ekman's seminal \citeyear{ekman:AMA1905} paper in combinations with
additional assumptions, such as a prescribed profile shape for eddy viscosity \cite{ellison:QJR1955} and two-layer models of the ABL take into account rotational effects at higher altitudes,
for instance when the wind speed needs to be evaluated at heights on the order of $100-200\,\mathrm{m}$,

@@

Still, in \citeyear{jiang:JAS2018}, \citeauthor{jiang:JAS2018} recognized that the outer
part of the Ekman boundary layer receives less attention in comparison with the surface layer
and study the neutral problem by Large-Eddy simulation (LES).
They focus on the wind speed and find an extended logarithmic layer when considering the wind
speed instead of the shear-aligned component, and they eventually demonstrate by means of
an analytical model that this vertical extension of the logarithmic layer may be explained
by a transfer of stress to the span-wise velocity component where it is assumed that the shear
vector $\tau(z)$ and stress vectors $(\p_zU,\p_zV)$ are aligned. 
More recently, \citet{ghannam:QJR2021} treated the horizontally averaged momentum budget to show
that departures from shear-alignment in the vicinity of the surface result in an integral of
the wind veer ($\alpha_M$ in their notation) over the height to very high accuracy ($\int_{z_0}^{H} \sin\alpha_M$ in their notation; their Eq.~(16)).

@@

homogeneous, stationary ABL over a flat surface under neutral stratification. 

@@

roughness or other physical complications encountered in the real geophysical system.

@@

relies on sub-grid closures that commonly assume alignment of the turbulent stress with gradients.
This pre-requisite is not fulfilled when the wind rotates with height.

@@

research on LES for the ABL over the past 50 years \citep{stoll:BM2020}. 

@@

of \emph{fully developed turbulence}  believed to occur for a sufficiently large but finite Reynolds number

@@

We consider here incompressible, turbulent Ekman flow, that is, the turbulent flow over a flat rotating plate, as a physical model for the truly neutral ABL.

@@

$\alpha(z)= \sphericalangle \left( \underline{G},\underline{U}(z)\right)$, where $\underline{G}=(G_1,G_2,0)$ is the geostrophic wind vector.

@@

that is governed by % the physical parameters
(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_2^2}$,

@@

The scales used in defining $\RE_\Lambda$ are of limited relevance for description of the turbulent flow state.
The turbulence scale separation in a wall-bounded flow is commonly characterized by the friction Reynolds number
\citep{jimenez:ARF2014}:

@@

@@

where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, $g_j=G_j/G$ are non-dimensionalized components geostrophic wind (with $g_{1}^2+g_{2}^2=1$ by construction), and $\epsilon$ is the Levi--Civita tensor.
These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral (streamwise and spanwise) directions, a no-slip--no-penetration boundary at $z=0$, and

@@

  \caption{\label{tab:sim-setup} Direct numerical simulation data sets used in this work. $Re_\Lambda$ and $Re_D$ refer to the Reynolds number defined in terms of the Rossby radius $\Lambda$ and Ekman-layer thickness $D$ respectively. $L_{xy}$ is the domain size in the stream- and span-wise direction. The grid is given by the number of grid points in the stream-wise ($N_x$), span-wise ($N_y$) and vertical ($N_z$) directions respectively. The resolution in the span-wise and stream-wise directions are given as $\Delta x^+$ and $\Delta y^+$. The grid in the vertical is stretched, and  resolution at the wall is given by $\Delta z^+$.}
  \centering\begin{tabular}{r r r c c c c c}

@@

       $125\,000$ &    $500$& 1.08 & $2048\times2048\times192$&4.1& 4.1& 1.05 \\

@@

       $281\,250$ & $   750$& 1.08 & $3072\times3072\times384$&5.6& 5.6& 1.60 \\ 

@@

    $1\,280\,000$ & $1\,600$& 0.54 & $3860\times7680\times960$&8.6& 4.3& 1.00\\
    \bottomrule 
  \end{tabular}

 \caption{\label{tab:data} DOIs and reference to the openly accesible data set at refubium repository} 
  \centering\begin{tabular}{r r r }
    \toprule
    $\RE_D$ & DOI & Reference \\
    \midrule
    \rowcolor{gray!30} 500& \DOI{10.17169/refubium-42505} & \cite{ansorge:2024c}\\
    \rowcolor{gray!30} 1000& \DOI{10.17169/refubium-42507}& \cite{ansorge:2024a}\\

    1300& \DOI{10.17169/refubium-42508} & \cite{ansorge:2024}  \\
    1600& \DOI{10.17169/refubium-42509} & \cite{ansorge:2024b} \\

@@

    normalized boundary layer depth $\delta_{95}/\delta$, 

@@

  \centering\begin{tabular}{r r r r r r r }

@@

Velocity profiles in inner units ($U^+(z^+)$, Fig.~\ref{fig:profiles}a)and outer units ($U^-(z^-)$, Fig.~\ref{fig:profiles}b)
are in accordance with previous work \citep{coleman:JFM1992,spalart:PF2008,spalart:PF2009,ansorge:BM2014,ansorge:BM2019}:

@@

geostrophic wind $G$) governs the inviscid normalization, i.e. a scaling independent of the Reynolds-number.

@@

No collapse in inner units is found for the profiles of spanwise velocity $V^{\alpha_\star+}(z^+)$. 

@@

    Circles in \textbf{(a)} and \textbf{(c)} denote the height
    $z^-=0.15$, diamonds in \textbf{(b)} and \textbf{(d)} are for $z^+=50$

@@

\end{figure} 
The organization of the flow is depicted in terms of the turbulence kinetic energy in Fig.~\ref{fig:slices}.
%
In vicinity of the wall, at $y^+\approx 10$, (Fig.~\ref{fig:slices}a), elongated streaks aligned with the
surface shear stress dominate.
%
Moving away from the wall, to $y^+\approx 150$ (well within the logarithmic region), the structures
are larger and more isotropic, but they are still largely aligned with the surface shear stress.
%
In the upper part of the outer layer, around $y^+\approx 1000$, no clear signature of the
surface veering direction is found, and intense TKE structures (bright yellow)
are organized on a large spatial scale with weaker eddies (greenish structures) and quiescent regions
in between.

@@

This precludes, first, a drag law wherewith we begin this section (\ref{sec:drag-law}).
Based on the scaling arguments put forward in Sec.~\ref{sec:scaling}, we then develop, second,
a formulation of the wind vector in the Ekman layer (Sec~\ref{sec:ekman}). Finally, we come up
with a separate formulation of the, third, stream-wise and, fourth, span-wise velocity components in the
overlap and inner regions of the flow. 

@@

  \label{eqn:zstar}

@@

\cite{optis:BM2014}, for instance, define an \emph{``effective geostrophic wind vector that has the same magnitude of the
observed surface geostrophic wind and is rotated by the angle $\alpha$ [their nomenclature]''} to overcome the

@@

The text-book solution for Ekman flow makes use of the physical boundary conditions (BCs) for
the~ABL (no-slip at the bottom and geostrophic wind in the free atmosphere) and a constant eddy

@@

        \end{array}\right)\right]. 

@@

with  $z_\mathrm{ek} = \delta_\mathrm{ek} (y^- -s_\mathrm{ek})$.

@@

the low-$\RE$ correction suggested by \cite{spalart:JFM1989}.

@@

Given the Ekman formulation of the velocity profile introduced in Sec.~\ref{sec:ekman},
one may express $v_{03}$ using the Ekman profile introduced in Sec.~\ref{sec:ekman} together
with the approximation for $u_\star(Re)$ found in Eq.~(\ref{eqn:zstar}). 

@@

\subsection{Implications for surface-layer scaling} 

@@

  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
  \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
    Thick, solid lines show theory, dashed lines data from DNS. }
  \label{fig:hodograph} 

\subsection{Comparison with other theories}

\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;

\textbf{Compare} to van Driest scaling

\textbf{Remember} of interpretation in the context of eddy viscosity (Fig.~\ref{fig:stresses}) 
Consider Townsend ``Turbulent Shear Flow''  Chapter 7.18: p. 319 
%
\par%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
$\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
%
\begin{figure}
  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
  \caption{Total velocity and veering (in degrees) vs inner and outer height.
    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
\end{figure} 
%
\section{Conclusions}
%
Applications:
\begin{itemize}
\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
\item
\end{itemize}

\bibliographystyle{spbasic_updated} 
\bibliography{ansorge}
